---
title: "Kevin Isaac Kareithi"
author: "Kevin Isaac Kareithi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Load necessary libraries
```{r code}
library(tidyverse) 
library(sf)
library(ggplot2)
library(ggmap)
library(plotly)
```
# Load the dataset
```{R data}
url <- "https://raw.githubusercontent.com/cema-uonbi/internship_task/main/data/cema_internship_task_2023.csv"
data <- read.csv(url)
head(data)
```
# Explore the structure and summary statistics of the dataset
```{R summary}
# Inspect data 
glimpse(data)
summary(data)
```
# Handling missing value
```{R missing}
# Fill in missing values for the columns with their mean
data <- data %>%
  mutate(
    Total.Dewormed = ifelse(is.na(Total.Dewormed), mean(Total.Dewormed, na.rm = TRUE), Total.Dewormed),
    Acute.Malnutrition = ifelse(is.na(Acute.Malnutrition), mean(Acute.Malnutrition, na.rm = TRUE), Acute.Malnutrition),
    stunted.6.23.months = ifelse(is.na(stunted.6.23.months), mean(stunted.6.23.months, na.rm = TRUE), stunted.6.23.months),
    stunted.0..6.months = ifelse(is.na(stunted.0..6.months), mean(stunted.0..6.months, na.rm = TRUE), stunted.0..6.months),
    stunted.24.59.months = ifelse(is.na(stunted.24.59.months), mean(stunted.24.59.months, na.rm = TRUE), stunted.24.59.months),
    diarrhoea.cases = ifelse(is.na(diarrhoea.cases), mean(diarrhoea.cases, na.rm = TRUE), diarrhoea.cases),
    Underweight.0..6.months = ifelse(is.na(Underweight.0..6.months), mean(Underweight.0..6.months, na.rm = TRUE), Underweight.0..6.months),
    Underweight.6.23.months = ifelse(is.na(Underweight.6.23.months), mean(Underweight.6.23.months, na.rm = TRUE), Underweight.6.23.months),
    Underweight.24.59.Months = ifelse(is.na(Underweight.24.59.Months), mean(Underweight.24.59.Months, na.rm = TRUE), Underweight.24.59.Months)
  )

# Inspect the filled data
glimpse(data)
```
# Transform to tidy format
```{R Tidy}
data_tidy <- data %>% 
  pivot_longer(c(Total.Dewormed,Acute.Malnutrition,stunted.6.23.months,stunted.0..6.months,stunted.24.59.months,diarrhoea.cases,Underweight.0..6.months,Underweight.6.23.months,Underweight.24.59.Months), 
               names_to = "variable", 
               values_to = "value")
```
# Visualize with histograms
# Create interactive histograms using plotly
```{R plot}
p <- ggplot(data_tidy, aes(x = value)) +
  geom_histogram() +
  facet_wrap(~variable, scales = "free") +
  theme_minimal()
ggplotly(p)
```

# Visualize with boxplots
# Create interactive boxplots using plotly
```{R boxplotl}
p <- ggplot(data_tidy, aes(x = period, y = value, color = variable)) + 
  geom_boxplot() +
  theme_minimal()+
  coord_flip()
ggplotly(p)
```

# Join with spatial data
```{R}
kenya_sf <- st_read("County.shp")
data$county <- gsub(" County$", "", data$county)
data_sf <- left_join(kenya_sf, data, by = c("Name" = "county"))
```
# Map acute malnutrition
# Create an interactive choropleth map using plotly

# State research question
# "What factors are associated with high rates of acute malnutrition across counties?" 

# Statistical analysis e.g. correlation, regression
# Calculate correlation between Acute Malnutrition and Dewormed
```{R correlation}
correlation <- cor(data_sf$`Acute.Malnutrition`, data_sf$Total.Dewormed)
print("Correlation coefficient between Acute Malnutrition and Dewormed:")
print(correlation)
```
# Fit a linear regression model to explore the relationship between Acute Malnutrition and Dewormed
```{R }
linear_model <- lm(Acute.Malnutrition ~ Total.Dewormed, data = data_sf)
summary(linear_model)
```
# Graphs showing the relationship between Acute Malnutrition and Dewormed
# Create interactive scatter plot with a regression line using plotly
```{R Graphs}
p <- ggplot(data_sf, aes(x = Total.Dewormed, y = Acute.Malnutrition)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Dewormed", y = "Acute Malnutrition") +
  theme_minimal()
ggplotly(p)
```

# Create interactive box plot using plotly
```{R box plot}
p <- ggplot(data_sf, aes(x = Total.Dewormed, y = Acute.Malnutrition)) +
  geom_boxplot() +
  labs(x = "Dewormed", y = "Acute Malnutrition") +
  theme_minimal()
ggplotly(p)
```